\plaintitle

### 2020-03-28 wip-design/paper.tex
Discussion and conclusions

% Evaluating ui systems research
% \cite{evaluating}

@@

\newpage

@@

\subsection{Open Journalism}
Journalism continually develops and responds to the many challenges it faces \cite{future}.
Two recent challenges are relevant to our work. The first is building trust in media.
One way of establishing trust in the age of fake news is to be more transparent about editorial
decisions, process and original sources. Many journalists believe that opening up the process
shows the quality and trustworthiness of their work~\cite{transparency}.
The second challenge is reader engagement. To develop a relationship with readers, journalists are
increasingly looking for meaningful ways of engagement. This includes reader comments, involvement
of citizen journalists \cite{comments,citizen} and the development of new interactive formats
\cite{youdraw}. To address the above challenges, a tool for data exploration should satisfy the
following three requirements.

@@

\newpage
~
\newpage

@@

~
\newpage

@@

\subsection{Conclusions}
The motivation behind The Gamma is to make simple programmatic data exploration accessible to
journalists and other non-experts. Main-stream programming languages have the benefit of
reproducibility and transparency, but are difficult to use. Our study shows that non-experts
are able to solve basic data exploration tasks using The Gamma. When asked whether The Gamma
is something that journalists could learn how to use, participant \#13 (a former journalist) answered:

@@

The Gamma is based on the iterative prompting principle which lowers the barrier to entry, but
it is worth noting that it does not fully eliminate complexity involved in data querying.
Instead, it provides a way of structuring the complexity. Iterative prompting makes it easy to
get started, but complex data sources will inevitably require additional learning. The Gamma
makes this easier by allowing users to learn from published code samples.
\newpage
~
\newpage

@@

\section{Discussion}

Data provenance (where data comes from)
What are the data sources availbale (how to start)?

\subsection{Study limitations}
exploratory in nature so we do not make any quantitative claims about effects

not comparing against other systems

\subsection{Design principles}
How well did we do wrt design principles?

\subsection{Design issues}
future challenges and limitations of the model - such as issues when modifying code
in the middle of the call chain

\section{Conclusions}

\newpage
~
\newpage

\section{Acknowledgments}
Yo

>>>>

@@

@@

\subsection{Open Journalism}
Journalism continually develops and responds to the many challenges it faces \cite{future}.
Two recent challenges are relevant to our work. The first is building trust in media.
One way of establishing trust in the age of fake news is to be more transparent about editorial
decisions, process and original sources.

Many journalists believe that opening up the process
shows the quality and trustworthiness of their work~\cite{transparency}.
The second challenge is reader engagement. To develop a relationship with readers, journalists are
increasingly looking for meaningful ways of engagement. This includes reader comments, involvement
of citizen journalists \cite{comments,citizen} and the development of new interactive formats
\cite{youdraw}. To address the above challenges, a tool for data exploration should satisfy the
following three requirements.

@@

@@

@@

\subsection{Study conclusions}
The motivation behind The Gamma is to make simple programmatic data exploration accessible to
journalists and other non-experts. Main-stream programming languages have the benefit of
reproducibility and transparency, but are difficult to use. Our study shows that non-experts
are able to solve basic data exploration tasks using The Gamma. When asked whether The Gamma
is something that journalists could learn how to use, participant \#13 (a former journalist) answered:

@@

The Gamma is based on the iterative prompting principle which lowers the barrier to entry, but
it is worth noting that it does not fully eliminate complexity involved in data querying.
Instead, it provides a way of structuring the complexity. Iterative prompting makes it easy to
get started, but complex data sources will inevitably require additional learning. The Gamma
makes this easier by allowing users to learn from published code samples.

@@

\section{Discussion}
We design a programming environment that can make programmatic data exploration accessible
to journalists. This is aligned with recent developments in journalism, which aims to build
trust through transparency and greater reproducibility as well as provide meaningful ways of
engagement.

\subsection{Evaluating The Gamma}
Data exploration environments are complex systems that do not yield to simple controlled
experimentation. We evaluate The Gamma using criteria inspired by Olsen~\cite{evaluating}.

\begin{itemize}
\item \emph{Importance.} Open journalism has the potential to make factual claims backed by
data more commonplace and enable wider audience engage with such claims. As such, we contribute
towards solving an important societal issue.

\item \emph{Expressive leverage.} The Gamma reduces the number of concepts that the user needs
to master to just the \emph{iterative prompting} principle, although complex data sources may
add more concepts through their domain specific languages.

\item \emph{Empowering new participants.} As demonstrated by our user study, The Gamma allows
non-experts, including those not comfortable with code, to perform basic programmatic
data exploration tasks.

\item \emph{Generality.} Finally, our case studies show that The Gamma can be used to solve
a wide range of data manipulation tasks involving many data sources including tabular data,
data cubes and graph databases.
\end{itemize}

\subsection{Remaining design issues}
There remain a number of aspects of data exploration in the context of journalism that
The Gamma does not address. Two of those, data provenance and data availability were also
observed by the participants in our study.

\paragraph{Data provenance}
Data sources such as \ikvd{olympics} or \ikvd{worldbank} are defined when initializing The Gamma
programming environment, but the system does not currently show where such data comes from.
For some data sources, the source is e.g.~a CSV file published by the government. In this case,
we can easily show the source. However, other type providers may pre-process data. Displaying
data source in such cases would require more sophisticated provenance tracking \cite{provenance}.

\paragraph{Data availability}
In the current version, The Gamma does not have a way of informing the user what data sources
are available. In other words, the user needs to know the first identifier, such as \ikvd{olympics},
to get started. We could address this by choosing a data source as the first step of iterative
prompting and perhaps typing \ikvd{.olympics}. However, a more fundamental issue is finding
the data source in the first place. This could be partly addressed by a type provider for a
currated online database such as Enigma Public\footnote{\url{https://docs.enigma.com/public}}.
Providing access to open government data repositories such as \href{http://data.gov.uk}{\small\bf\ttfamily data.gov.uk}
is more appealing, but very challenging due to their unstructured nature.

\subsection{Evaluation limitations}
The evaluation conducted in this paper is qualitative and exploratory in nature. In particular,
we do not make any quantitative claims about the usability of The Gamma and its learning curve.
Our study also focused on the core iterative prompting interaction, but our case studies often
required using other features such as operations with parameters.

Although The Gamma is open-source, it has not been deployed in production in a newsroom so far.
This would lead to valuable insights, but it requires finding a suitable fortuitous opportunity.
Finally, we also do not compare the usability of The Gamma with the usability of other popular
systems such as Tableau \cite{tableau}. It would be possible to set tasks that can be completed
in both systems, but the systems have very different aims making such comparison problematic.

\section{Conclusions}
We present The Gamma, a simple data exploration environment. The Gamma is motivated by the needs
of journalists working with data, who need to produce transparent and easy-to-understand reports
backed by data that are transparent and engaging, all the while facing tight deadlines.

The Gamma is based on a single interaction principle, \emph{iterative prompting}, can be used to
complete a wide range of data exploration tasks using a variety of data sources including tabular
data, data cubes and graph databases. The design lowers the barrier to entry for programmatic
data exploration and it makes it easy to learn the system independently from examples and by
experimentation.

% \section{Acknowledgments}
% Yo
